---
title: "PA1_template.Rmd"
author: "James Thomson"
date: "September 29, 2017"
output: html_document
---
# Loading the data

```{r}
data <- read.csv("C:/Users/jthomson/Documents/activity.csv")
```

# What is mean total number of steps taken per day?

```{r}
stepsdate <- aggregate(steps ~ date, data, sum, na.rm = TRUE)
```

#### Mean and median number of steps taken each day:

```{r}
mean(stepsdate$steps)
median(stepsdate$steps)
```

#### Histogram of the total number of steps taken each day

```{r}
plot(stepsdate$steps,
        type = "h",
        main = "Total Number of Steps Taken Each Day",
        xlab = "Date", 
        ylab = "Steps",
        col = "red")
```

# What is the average daily activity pattern?

```{r}
stepsint <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)

plot(stepsint,
        type = "l",
        main = "Average Daily Activity Pattern",
        xlab = "Interval (5-minute)",
        ylab = "Average Number of Steps",
        col = "red")
```

#### The 5-minute interval that, on average, contains the maximum number of steps

```{r}
stepsint$interval[which.max(stepsint$steps)]
```

# Imputing missing values

#### Count of missing (NA) data

```{r}
sum(is.na(data))
```

#### Replace missing data with 0
```{r}
imputed.data <- data
imputed.data$steps[is.na(imputed.data$steps)] <- 0
```

#### Mean and median number of steps taken each day:

```{r}
mean(imputed.data$steps)
median(imputed.data$steps)
```

#### Histogram of the total number of steps taken each day

```{r}
plot(imputed.data$steps, 
        type = "h",
        main = "Total Number of Steps Taken Each Day",
        xlab = "Date", 
        ylab = "Steps",
        col = "red") 
```

# Are there differences in activity patterns between weekdays and weekends?

#### Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed.data$day <- as.factor(ifelse(is.element(weekdays(as.Date(imputed.data$date)),weekdays), "Weekday", "Weekend"))

steps.interval <- aggregate(steps ~ interval + day, imputed.data, mean)

library(lattice)

xyplot(steps.interval$steps ~ steps.interval$interval|steps.interval$day,
        type="l",
        main="Total Number of Steps Taken Each Day",
        xlab="Interval (5-minute)", 
        ylab="Steps",
        col = "red",
        layout=c(1,2))